Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars

### 2014-10-19 drafts/cfp/draft.tex
Draft changes.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{anyfontsize}
\usepackage{subcaption}
\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
\author{Eyal Dechter \& Josh Rule \& Josh Tenenbaum}

@@

\section{Abstract}
A popular approach in statistical machine learning is to automatically
find the latent \emph{features} of the data that are most relevant or
informative. But a latent feature based view of concept learning does
not match up naturally with the structured symbolic knowledge that
seems to underlie people's understanding of many domains. How can we
combine the flexibility, noise tolerance, and computational
practicality of latent feature based statistical machine learning with
the symbolic knowledge representations that best capture human
understanding? Here present a learning model we dub Latent Concept
Grammars (LCGs) as a proposal. An LCG is a densely connected set of
probabilistic string rewrite rules over an initially uninterpreted set
of predicates. When fit to data using a hierarchical Bayesian prior,
these predicates assume the roles of informative latent concepts. This
approach allows us to learn knowledge that is stochastic and
recursively structured while maintaining computational
tractability. We show how learning in LCGs can be implemented using
the probabilistic logic programming system PRISM, and we present
preliminary experiments.

How do people learn systems of related concepts from
data?  Consider, for example, a school-age child's knowledge of
number. This knowledge supports a large number of related operations:
the child knows which words denote numbers, how to use those words to
label cardinalities of sets, how to translate between digit
representations and word representations of number, how to compare
cardinalities, etc. The child knows how to perform these operations
over an unbounded set of elements, even though the data that underlies
this knowledge is small and noisy. Here we propose that such systems
of related concepts can be learned via hierarchical probabilistic
inference over probabilistic string rewrite systems (PSRS) -- a mildly
context-sensitive class of probabilistic string gramamrs that
corresponds to a restricted fragment of stochastic logic programs
(SLPs). Specifically, we extend the variational
expectation-maximization algorithm for Hierarchical Dirichlet PCFGs to
accomodate PSRSs. This algorithm we present enables a tractible
compromise between highly restrictive context-free knowledge
representations and the intractability of performing similar inference
procedures over general logic programs. Empirically, we show how such
restricted string rewrite systems can capture non-trivial conceptual
knowledge and how our algorithm is able to learn such knowledge from
data.

\section{Abstract 2}
What does it mean to understand a concept, and how do people come to
such an understanding.

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{anyfontsize}
\usepackage{subcaption}

\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
\author{Eyal Dechter \& Josh Rule \& Josh Tenenbaum}

@@

\section{Abstract}
To what extent can the meaning of language be learned from linguistic
data alone?  The sentences that people say contain a great deal of
information about the meaning of linguistic concepts. Most approaches
to semantics assume an explicit knowledge representation formalism
that ties utterances to propositions about things in the world. The
study of linguistic syntax, however, has largely assumed that the
meaning of syntactic structures is encoded directly into the grammar
of a language. For machine learning, extending the grammatical
embedding of meaning to at least some of the linguistic concepts is
attractive because it reduces the problem of automated language
learning to the problem of learning a distribution over naturally
occuring utterances. To what extent is this a feasible approach? Here,
we present a first attempt at the automated induction of probabilistic
context sensitive grammars which we call Latent Predicate Networks
(LPNs), motivated by the idea that the predicates and rules that are
learned correspond to the concepts expressed by utterances. We show
that for a small fragment of sentences expressing relationships
between English number words, we can use hierarchical Bayesian
inference to learn LPNs that can answer simple queries about
previously unseen linguistic concepts within this domain. These
generalizations, we argue, demonstrate LPNs' promise as a tool for
learning and representing conceptual knowledge in language.

% A popular approach in statistical machine learning is to automatically
% find the latent \emph{features} of the data that are most relevant or
% informative. But a latent feature based view of concept learning does
% not match up naturally with the structured symbolic knowledge that
% seems to underlie people's understanding of many domains. How can we
% combine the flexibility, noise tolerance, and computational
% practicality of latent feature based statistical machine learning with
% the symbolic knowledge representations that best capture human
% understanding? Here present a learning model we dub Latent Concept
% Grammars (LCGs) as a proposal. An LCG is a densely connected set of
% probabilistic string rewrite rules over an initially uninterpreted set
% of predicates. When fit to data using a hierarchical Bayesian prior,
% these predicates assume the roles of informative latent concepts. This
% approach allows us to learn knowledge that is stochastic and
% recursively structured while maintaining computational
% tractability. We show how learning in LCGs can be implemented using
% the probabilistic logic programming system PRISM, and we present
% preliminary experiments.

% How do people learn systems of related concepts from
% data?  Consider, for example, a school-age child's knowledge of
% number. This knowledge supports a large number of related operations:
% the child knows which words denote numbers, how to use those words to
% label cardinalities of sets, how to translate between digit
% representations and word representations of number, how to compare
% cardinalities, etc. The child knows how to perform these operations
% over an unbounded set of elements, even though the data that underlies
% this knowledge is small and noisy. Here we propose that such systems
% of related concepts can be learned via hierarchical probabilistic
% inference over probabilistic string rewrite systems (PSRS) -- a mildly
% context-sensitive class of probabilistic string gramamrs that
% corresponds to a restricted fragment of stochastic logic programs
% (SLPs). Specifically, we extend the variational
% expectation-maximization algorithm for Hierarchical Dirichlet PCFGs to
% accomodate PSRSs. This algorithm we present enables a tractible
% compromise between highly restrictive context-free knowledge
% representations and the intractability of performing similar inference
% procedures over general logic programs. Empirically, we show how such
% restricted string rewrite systems can capture non-trivial conceptual
% knowledge and how our algorithm is able to learn such knowledge from
% data.

% \section{Abstract 2}
% What does it mean to understand a concept, and how do people come to
% such an understanding.
